Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-25 numerical-mhd-dedner.tex
Many small changes to all files -- GB final edits

To more clearly describe the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics in the form of conservation laws as,
\begin{equation}
 \frac{\partial{U}}{\partial{t}} +
 \frac{\partial{F^x}}{\partial{x}} + \frac{\partial{F^y}}{\partial{y}} + \frac{\partial{F^z}}{\partial{z}}= 0, \label{hydro}
\end{equation}
The conserved variable $U$ is given by
\begin{equation}
 U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E)^{T},
\end{equation} 
where $\rho$ is density, $v_i$ are the three components of velocity
for $i={x,y,z}$, $E=v^2/2 + e$ denotes the specific total energy and $e$ the
specific internal energy.

For the generalized Lagrange multiplier formulation of the MHD
equations \citep{2002JCoPh.175..645D}, we consider these 
conserved variables
\begin{equation}
 U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E+B^2/2, B_x, B_y, B_z, \psi)^{T},
\end{equation} 
where $B_i$ with $i={x,y,z}$ are the three components of magnetic
fields and $\psi$ is the additional scalar field introduced in the GLM
formulation for the divergence cleaning.  The fluxes then are
\begin{eqnarray}
 F^x &=& (\rho v_x, \rho v_x^2+p+B^2/2-B_x^2, \rho v_yv_x-B_yB_x, \cr
 && \rho v_zv_x-B_zB_x, \rho ({v^2\over2} + h)v_x+B^2v_x-B_xB\cdot v, \cr
&& \psi, v_xB_y-v_yB_x, -v_zB_x+v_xB_z, c_h^2B_x)^T,\\
 F^y &=& (\rho v_y, \rho v_xv_y-B_xB_y, \rho v_y^2+p+B^2/2-B_y^2, \cr
 && \rho v_zv_y-B_zB_y, \rho ({v^2\over2} + h)v_y+B^2v_y-B_yB\cdot v, \cr
 && v_yB_z-v_zB_y,\psi,-v_xB_y+v_yB_x, c_h^2B_y)^{T}, \\
 F^z &=& (\rho v_z, \rho v_xv_z-B_xB_z, \rho v_yv_z-B_yB_z, \rho v_z^2+p+B^2/2-B_z^2, \cr
 && \rho ({v^2\over2} + h)v_z+B^2v_z-B_zB\cdot v, \cr
    &&  -v_yB_z+v_zB_y, v_zB_x-v_xB_z,\psi, c_h^2B_z)^{T},
\end{eqnarray}
where $c_h$ is a constant controlling the propagation speed and
damping rate of $\div B$, and $h=e+p/\rho$ denotes the enthalpy.
All quantities are cell-centered.

@@

Finally, we note that for cosmological simulations, this solver uses a
slightly different definition of the magnetic field than used in the
rest of the paper.  In particular, the field is defined as $\vecB =
a^{3/2} \vecB^{\prime}$ (where $\vecB^{\prime}$ is the proper field
strength).  This adds a source term of $-\dot{a}/(2a) \vecB$ on the
right-hand side of Equation~\ref{eq:induction} and removes the factor
of $a$ in the $\vecB$ term in the energy equation
(Eq.~\ref{eq:total_energy_def}) and in the definition of the isotropic
pressure $p^*$.

>>>>

To more clearly describe the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics in the form of conservation laws as,
\begin{equation}
 \frac{\partial{U}}{\partial{t}} +
 \frac{\partial{F^x}}{\partial{x}} + \frac{\partial{F^y}}{\partial{y}} + \frac{\partial{F^z}}{\partial{z}}= 0, \label{hydro}
\end{equation}
The conserved variable $U$ is given by
\begin{equation}
 U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E)^{T},
\end{equation} 
where $\rho$ is density, $v_i$ are the three components of velocity
for $i={x,y,z}$, $E=v^2/2 + e$ denotes the specific total energy and $e$ the
specific internal energy (note that in this section only, we use the specific
energy).

For the generalized Lagrange multiplier (GLM) formulation of the MHD
equations \citep{2002JCoPh.175..645D}, we consider these 
conserved variables
\begin{equation}
 U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E+B^2/2, B_x, B_y, B_z, \psi)^{T},
\end{equation} 
where $B_i$ with $i={x,y,z}$ are the three components of magnetic
fields and $\psi$ is the additional scalar field introduced in the GLM
formulation for the divergence cleaning.  The fluxes then are
\begin{eqnarray}
 F^x &=& (\rho v_x, \rho v_x^2+p+B^2/2-B_x^2, \rho v_yv_x-B_yB_x, \cr
 && \rho v_zv_x-B_zB_x, \rho ({v^2\over2} + h)v_x+B^2v_x-B_xB\cdot v, \cr
&& \psi, v_xB_y-v_yB_x, -v_zB_x+v_xB_z, c_h^2B_x)^T,\\
 F^y &=& (\rho v_y, \rho v_xv_y-B_xB_y, \rho v_y^2+p+B^2/2-B_y^2, \cr
 && \rho v_zv_y-B_zB_y, \rho ({v^2\over2} + h)v_y+B^2v_y-B_yB\cdot v, \cr
 && v_yB_z-v_zB_y,\psi,-v_xB_y+v_yB_x, c_h^2B_y)^{T}, \\
 F^z &=& (\rho v_z, \rho v_xv_z-B_xB_z, \rho v_yv_z-B_yB_z, \rho v_z^2+p+B^2/2-B_z^2, \cr
 && \rho ({v^2\over2} + h)v_z+B^2v_z-B_zB\cdot v, \cr
    &&  -v_yB_z+v_zB_y, v_zB_x-v_xB_z,\psi, c_h^2B_z)^{T},
\end{eqnarray}
where $c_h$ is a constant controlling the propagation speed and
damping rate of $\div B$, and $h=e+p/\rho$ denotes the specific 
enthalpy.  All quantities are cell-centered.

@@

Finally, we note that for cosmological simulations, this solver uses a
slightly different definition of the magnetic field than used in the
rest of the paper.  In particular, the field is defined as $\vecB =
a^{3/2} \vecB^{\prime}$ (where $\vecB^{\prime}$ is the proper field
strength).  This adds a source term of $-\dot{a}/(2a) \vecB$ on the
right-hand side of Equation~(\ref{eq:induction}) and removes the factor
of $a$ in the $\vecB$ term in the energy equation
(Eq.~\ref{eq:total_energy_def}) and in the definition of the isotropic
pressure $p^*$.

### 2013-06-25 numerical-ppm.tex
Many small changes to all files -- GB final edits

The remainder of the terms in the fluid equations
involve derivatives.  
Because we are interested in phenomena with no special geometry, we will
restrict discussion to Cartesian coordinates.
We can dimensionally split the equations and rewrite the one-dimensional (Eulerian) 
versions of equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
in conservative form as,
\begin{eqnarray}
\frac{\partial \rho}{\partial t}  + \frac{1}{a} \frac{\partial \rho v }{\partial x}    & = &  
     0 \label{eq:mass1d} \\
\frac{\partial \rho v}{\partial t}  + \frac{1}{a} \frac{\partial \rho v^2}{\partial x}   + 
      \frac{1}{a} \frac{\partial p}{ \partial x} & = & 
      \rho \frac{1}{a} \frac{\partial \phi}{ \partial x}  \label{eq:momentum1d}  \\
\frac{\partial \rho E}{\partial t}  + \frac{1}{a} \frac{\partial \rho v E}{\partial x}  & =  &
      \rho v \frac{1}{a} \frac{\partial \phi}{\partial x}. \label{eq:total_energy1d}
\end{eqnarray}
%
Here $x$ and $v$ refer to the one dimensional comoving position and 
peculiar velocity of the baryonic gas, and $g$ is the acceleration at the cell center.
These equations are now in a form that can be solved by the split PPM scheme.

@@

We note that the acceleration required in Equation~(\ref{eq:momentum_diff}) is actually the
acceleration felt by the entire zone and not just at the zone center.
Therefore, it is possible to find the mass-weighted average acceleration over the zone
by expanding the density and acceleration distributions and
retaining all terms up to second-order in $\Delta x$:
%
\begin{equation}
g_j^{n+1/2} = 
       \frac{1}{2 a^{n+1/2} \delta x_j} \left[ 
             \phi_{j+1}^{n+1/2} 
           - \phi_{j-1}^{n+1/2} 
           + \frac{1}{12} \left(    \phi^{n+1/2}_{j+1} 
                                 - 2\phi^{n+1/2}_j 
                                 + \phi^{n+1/2}_{j-1} \right) 
                                   {\delta d_j \over d_j}
       \right].
       \label{eq:accel}
\end{equation}
Generally we find that the potential is so slowly varying that this change is unnecessary.

@@

In this section, we describe the methods we adopt to solving the Riemann
problem, which is generally required to compute the fluxes in 
any Godunov-based scheme.  This section therefore applies to all three of our 
Godunov-based schemes.  The Riemann problem we are solving
involves two constant states separated by a single discontinuity at $t=0$.
The subsequent evolution has an exact analytic solution.  This solution is
described in detail in many texts on computational fluid dynamics
\citep[e.g.,][]{toro-1997}.   In brief, there are three waves that propagate
away from the initial discontinuity.  The central wave, characterized by a
density jump but not a pressure jump, is called the contact discontinuity.  The
waves traveling to the left and right of the contact discontinuity can be either
shocks, if characteristics converge on the wave front, or rarefaction fans if
characteristics diverge.

While there exists an exact solution to this problem, finding it is expensive.  There are four
possible combination of left- and right- traveling shocks and rarefactions,
only one of which is fully consistent with the initial conditions.  Once the
correct physical state is determined, the pressure in the central region can
only be determined by finding the root to an algebraic equation, which is necessarily
an iterative process.  Thus a series of \emph{approximate} Riemann solvers are
typically used.  There are four approximate Riemann solvers in \enzo: two-shock
\citep{toro-1997},
Harten-Lax-van Leer \citep[HLL,][]{toro-1997}, HLL with a contact discontinuity
 \citep[HLLC,][]{toro-1997}, and HLL with
multiple discontinuities \citep[HLLD,][]{Miyoshi05}.  Two-shock is used only
with the PPM method.  HLL and HLLC are used with PPM, MUSCL (both with and
without MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works
with both the
MUSCL and MHD-CT methods.

>>>>

The remainder of the terms in the fluid equations
involve derivatives.  
Because we are interested in phenomena with no special geometry, we will
restrict discussion to Cartesian coordinates.
We can dimensionally split the equations and rewrite the one-dimensional (Eulerian) 
versions of equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
in conservative form as,
\begin{eqnarray}
\frac{\partial \rho}{\partial t}  + \frac{1}{a} \frac{\partial \rho v }{\partial x}    & = &  
     0 \label{eq:mass1d} \\
\frac{\partial \rho v}{\partial t}  + \frac{1}{a} \frac{\partial \rho v^2}{\partial x}   + 
      \frac{1}{a} \frac{\partial p}{ \partial x} & = & 
      \rho \frac{g}{a}  \label{eq:momentum1d}  \\
\frac{\partial \rho E}{\partial t}  + \frac{1}{a} \frac{\partial \rho v E}{\partial x}  & =  &
      \rho v \frac{g}{a} \label{eq:total_energy1d}
\end{eqnarray}
%
Here $x$ and $v$ refer to the one dimensional comoving position and 
peculiar velocity of the baryonic gas, and $g$ is the acceleration at the cell center.
These equations are now in a form that can be solved by the split PPM scheme.

@@

We note that the acceleration required in Equation~(\ref{eq:momentum_diff}) is actually the
acceleration felt by the entire zone and not just at the zone center.
Therefore, it is possible to find the mass-weighted average acceleration over the zone
by expanding the density and acceleration distributions and
retaining all terms up to second-order in $\Delta x$ \citep{1995CoPhC..89..149B}, although we find that
the potential is so slowly varying that this is unnecessary.
%
%\begin{equation}
%g_j^{n+1/2} = 
%       \frac{1}{2 a^{n+1/2} \delta x_j} \left[ 
%             \phi_{j+1}^{n+1/2} 
%           - \phi_{j-1}^{n+1/2} 
%           + \frac{1}{12} \left(    \phi^{n+1/2}_{j+1} 
%                                 - 2\phi^{n+1/2}_j 
%                                 + \phi^{n+1/2}_{j-1} \right) 
%                                   {\delta d_j \over d_j}
%       \right].
%       \label{eq:accel}
%\end{equation}
%Generally we find that the potential is so slowly varying that this change is unnecessary.

@@

In this section, we describe the methods we adopt to solve the Riemann
problem, which is generally required to compute the fluxes in 
any Godunov-based scheme.  This section therefore applies to all three of our 
Godunov-based schemes.  The Riemann problem we are solving
involves two constant states separated by a single discontinuity at $t=0$.
The subsequent evolution has an exact analytic solution.  This solution is
described in detail in many texts on computational fluid dynamics
\citep[e.g.,][]{toro-1997}.   In brief, there are three waves that propagate
away from the initial discontinuity.  The central wave, characterized by a
density jump but not a pressure jump, is called the contact discontinuity.  The
waves traveling to the left and right of the contact discontinuity can be either
shocks, if characteristics converge on the wave front, or rarefaction fans if
characteristics diverge.

While there exists an exact solution to this problem, finding it is expensive.  There are four
possible combination of left- and right- traveling shocks and rarefactions,
only one of which is fully consistent with the initial conditions.  Once the
correct physical state is determined, the pressure in the central region can
only be computed by finding the root to an algebraic equation, which is necessarily
an iterative process.  Thus a series of \emph{approximate} Riemann solvers are
typically used.  There are four approximate Riemann solvers in \enzo: two-shock
\citep{toro-1997},
Harten-Lax-van Leer \citep[HLL,][]{toro-1997}, HLL with a contact discontinuity
 \citep[HLLC,][]{toro-1997}, and HLL with
multiple discontinuities \citep[HLLD,][]{Miyoshi05}.  Two-shock is used only
with the PPM method.  HLL and HLLC are used with PPM, MUSCL (both with and
without MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works
with both the
MUSCL and MHD-CT methods.

### 2013-06-25 numerical-radiation-raytracing.tex
added renyue cen changes.
Many small changes to all files -- GB final edits

The rays are traced through the grid in a typical fashion
\citep[e.g.][]{Abel99_RT}, in which we calculate the next cell
boundary crossing.  The ray segment length crossing the cell is
%
\begin{equation}
  \label{eqn:trace}
  dr = R_0 - \min_{i=1 \rightarrow 3} \left[(x_{\rm cell,i} - x_{\rm src,i}) /
    \hat{n}_{\rm ray,i} \right],
\end{equation}
%
where $R_0$, $\hat{n}_{\rm ray}$, $x_{\rm cell,i}$, and $x_{\rm
  src,i}$ are the initial distance traveled by the ray, normal
direction of the ray, the next cell boundary crossing in the $i$-th
dimension, and the position of the point source that emitted the ray,
respectively.  However before the ray travels across the cell, we
evaluate the ratio of the face area $A_{\rm cell}$ of the current cell
and the solid angle $\Omega_{\rm ray}$ of the ray,
%
\begin{equation}
  \label{eqn:split}
  \Phi_c = \frac{A_{\rm cell}} {\Omega_{\rm ray}} = 
  \frac{N_{\rm pix} (\Delta x)^2} {4\pi R_0^2}.
\end{equation}
%
If $\Phi_c$ is less than a pre-determined value (usually $>3$), the
ray is split into 4 child rays.  The pixel numbers of the child rays
$p^\prime$ are given by the ``nested'' scheme of HEALPix at the next
level, i.e. $p^\prime = 4 \times p + [0,1,2,3]$, where $p$ is the
original pixel number.  The child rays (1) acquire the new normal
vectors of the pixels, (2) retain the same radius of the parent ray,
and (3) gets a quarter of the photon flux of the parent ray.
Afterward, the parent ray is discontinued.

@@

The radiation field is calculated by integrating Equation
(\ref{eqn:rtray}) along each ray, which is done by considering the
discretization of the ray into segments.  In the following
description, we assume the rays are monochromatic for simplicity.  For
convenience, we express the integration in terms of optical depth
$\tau = \int \kappa(r,t) \; dr$, and for a ray segment,
%
\begin{equation}
  \label{eqn:dtau}
  d\tau = \sigma_{\rm abs}(\nu) n_{\rm abs} dr.
\end{equation}
Here $\sigma_{\rm abs}$ and $n_{\rm abs}$ are the cross section and
number density of the absorbing medium, respectively.  In the static
case, equation (\ref{eqn:rtray}) has a simple exponential analytic
solution, and the photon flux of a ray is reduced by
%
\begin{equation}
  \label{eqn:flux}
  dP = P \times (1 - e^{-\tau})
\end{equation}
as it crosses a cell.  We equate the photo-ionization rate to the
absorption rate, resulting in photon conservation \citep{Abel99_RT,
  Mellema06}.  Thus the photo-ionization and
photo-heating rates associated with a single ray ($k_{\rm ph}$ and
$\Gamma_{\rm ph}$, respectively)  are
%
\begin{equation}
  \label{eqn:kph}
  k_{\rm ph} = \frac{P (1 - e^{-\tau})}{n_{\rm abs} \; V_{\rm cell} \; dt_P},
\end{equation}
\begin{equation}
  \label{eqn:gamma}
  \Gamma_{\rm ph} = k_{\rm ph} \; (E_{\rm ph} - E_i),
\end{equation}
where $V_{\rm cell}$ is the cell volume, $E_{\rm ph}$ is the photon
energy, and $E_i$ is the ionization energy of the absorbing material.
In each cell, the photo-ionization and photo-heating rates from each
ray in the calculation are summed. After the ray tracing is
complete, these rates are used as inputs to the solver described in
Section~\ref{sec.ov.chem} to update the ionization, chemical, and energy states
of the gas in each cell.

>>>>

The rays are traced through the grid in a typical fashion
\citep[e.g.][]{Abel99_RT}, in which we calculate the next cell
boundary crossing.  The ray segment length crossing the cell is
%
\begin{equation}
  \label{eqn:trace}
  dr = R_0 - \min_{i=1 \rightarrow 3} \left[(x_{\rm cell,i} - x_{\rm src,i}) /
    \hat{n}_{\rm ray,i} \right],
\end{equation}
%
where $R_0$, $\hat{n}_{\rm ray}$, $x_{\rm cell,i}$, and $x_{\rm
  src,i}$ are the initial distance traveled by the ray, normal
direction of the ray, the next cell boundary crossing in the $i$-th
dimension, and the position of the point source that emitted the ray,
respectively.  However before the ray travels across the cell, we
evaluate the ratio of the face area $A_{\rm cell}$ of the current cell
and the solid angle $\Omega_{\rm ray}$ of the ray,
%
\begin{equation}
  \label{eqn:split}
  \Phi_c = \frac{A_{\rm cell}} {\Omega_{\rm ray}} = 
  \frac{N_{\rm pix} (\Delta x)^2} {4\pi R_0^2}.
\end{equation}
%
If $\Phi_c$ is less than a pre-determined value (usually $>3$), the
ray is split into 4 child rays.  The pixel numbers of the child rays
$p^\prime$ are given by the ``nested'' scheme of HEALPix at the next
level, i.e. $p^\prime = 4 \times p + [0,1,2,3]$, where $p$ is the
original pixel number.  The child rays (1) acquire the new normal
vectors of the pixels, (2) retain the same radius of the parent ray,
and (3) get a quarter of the photon flux of the parent ray.
Afterward, the parent ray is discontinued.

@@

The radiation field is calculated by integrating Equation
(\ref{eqn:rtray}) along each ray, which is done by considering the
discretization of the ray into segments.  In the following
description, we assume the rays are monochromatic for simplicity.  For
convenience, we express the integration in terms of optical depth
$\tau = \int \kappa(r,t) \; dr$, and for a ray segment
%
\begin{equation}
  \label{eqn:dtau}
  d\tau = \sigma_{\rm abs}(\nu) n_{\rm abs} dr.
\end{equation}
Here $\sigma_{\rm abs}$ and $n_{\rm abs}$ are the cross section and
number density of the absorbing medium, respectively.  In the static
case, equation (\ref{eqn:rtray}) has a simple exponential analytic
solution, and the photon flux of a ray is reduced by
%
\begin{equation}
  \label{eqn:flux}
  dP = P \times (1 - e^{-\tau})
\end{equation}
as it crosses a cell.  We equate the photo-ionization rate to the
absorption rate, resulting in photon conservation \citep{Abel99_RT,
  Mellema06}.  Thus the photo-ionization and
photo-heating rates associated with a single ray ($k_{\rm ph}$ and
$\Gamma_{\rm ph}$, respectively)  are
%
\begin{equation}
  \label{eqn:kph}
  k_{\rm ph} = \frac{P (1 - e^{-\tau})}{n_{\rm abs} \; V_{\rm cell} \; dt_P},
\end{equation}
\begin{equation}
  \label{eqn:gamma}
  \Gamma_{\rm ph} = k_{\rm ph} \; (E_{\rm ph} - E_i),
\end{equation}
where $V_{\rm cell}$ is the cell volume, $E_{\rm ph}$ is the photon
energy, and $E_i$ is the ionization energy of the absorbing material.
In each cell, the photo-ionization and photo-heating rates from each
ray in the calculation are summed. After the ray tracing is
complete, these rates are used as inputs to the solver described in
Section~\ref{sec.ov.chem} to update the ionization, chemical, and energy states
of the gas in each cell.

### 2013-06-25 numerical-starformation.tex
Many small changes to all files -- GB final edits

Due to the computationally unfeasible number of stars in a galaxy
($10^{11}$) and the lack of detailed understanding of star formation, a
number of phenomenological star formation models are included in
\enzo.
%The \enzo\ code includes approximately ten models for star formation
%and feedback.  
Broadly speaking, these methods all work in similar
ways: At a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., that have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but can also include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with
the same mass is placed in the center of that cell with the same
velocity as the removed gas.  This star particle is then
allowed to inject mass, momentum, thermal energy, metals, and possibly
magnetic fields and/or cosmic ray populations into its local
environment.  In general, the particle is treated as an ensemble of
stars, with feedback properties occuring over time according to the
assumed initial mass function of the stellar population.

@@

The methods described in the previous two sections are generally used
in simulations that have very poor resolution, $\Delta x \ga 1$~kpc.
At this physical scale, individual star forming regions are not
resolved, so all uncertainty about the behavior of molecular clouds is
folded into a density threshold for star formation.  In calculations
with much higher resolution, however (on the order of a few pc), individual
molecular clouds can be resolved, thus rendering these approximations
invalid.  To that end, \citet{2012ApJ...749...36K} have implemented
a star formation algorithm that is specifically geared to
high-resolution cosmological simulations of galaxy formation, where
the formation of molecular hydrogen is followed directly and stars
are allowed to form at the highest level of refinement when the local
H$_2$ fraction exceeds a pre-determined threshold.  Cells are examined
every root grid timestep, $\Delta t_0$, and cells at the highest level
that exceed the H$_2$ limit form star particles with masses
proportional to the inferred mass of molecular hydrogen in the
star-forming region
\citep{2008ApJ...689..865K,2009ApJ...693..216K,2010ApJ...709..308M}.
The mass of the particle is calculated as:
\begin{equation} 
m_p = \epsilon \rho_{\rm gas} (\Delta x_{m})^3 \frac{\Delta t_0}{t_*}
\end{equation}
where $\Delta x_{m}$ is the resolution of the maximum level of
refinement, $\epsilon$ is an efficiency parameter with a standard
value of 0.01 \citep[as motivated by][]{2007ApJ...654..304K}, and
t$_*$ is the local free-fall timescale.

>>>>

Due to the computationally unfeasible number of stars in a galaxy
($10^{11}$) and the lack of detailed understanding of star formation, a
number of phenomenological star formation models are included in
\enzo.
%The \enzo\ code includes approximately ten models for star formation
%and feedback.  
Broadly speaking, these methods all work in similar
ways: at a specified time interval, all grid cells that are at the
highest local level of refinement (i.e., that have no child cells) are
examined to see if they meet a set of criteria for star formation.
This may simply be a baryon density threshold, but can also include more
complex tests, such as an examination of local cooling and dynamical
time scales, molecular hydrogen fraction, metallicity, and converging
gas flows.  If the star formation criteria are met, some mass of gas
is taken away from the cell in question and a ``star particle'' with
the same mass is placed in the center of that cell with the same
velocity as the removed gas.  This star particle is then
allowed to inject mass, momentum, thermal energy, metals, and possibly
magnetic fields and/or cosmic ray populations into its local
environment.  In general, the particle is treated as an ensemble of
stars, with feedback properties occuring over time according to the
assumed initial mass function of the stellar population.

@@

The methods described in the previous two sections are generally used
in simulations that have relatively poor resolution, $\Delta x \ga 1$~kpc.
At this physical scale, individual star forming regions are not
resolved, so all uncertainty about the behavior of molecular clouds is
folded into a density threshold for star formation.  In calculations
with much higher resolution, however (on the order of a few pc), individual
molecular clouds can be resolved, thus rendering these approximations
invalid.  To that end, \citet{2012ApJ...749...36K} have implemented
a star formation algorithm that is specifically geared to
high-resolution cosmological simulations of galaxy formation, where
the formation of molecular hydrogen is followed directly and stars
are allowed to form at the highest level of refinement when the local
H$_2$ fraction exceeds a pre-determined threshold.  Cells are examined
every root grid timestep, $\Delta t_0$, and cells at the highest level
that exceed the H$_2$ limit form star particles with masses
proportional to the inferred mass of molecular hydrogen in the
star-forming region
\citep{2008ApJ...689..865K,2009ApJ...693..216K,2010ApJ...709..308M}.
The mass of the particle is calculated as:
\begin{equation} 
m_p = \epsilon \rho_{\rm gas} (\Delta x_{m})^3 \frac{\Delta t_0}{t_*}
\end{equation}
where $\Delta x_{m}$ is the resolution of the maximum level of
refinement, $\epsilon$ is an efficiency parameter with a standard
value of 0.01 \citep[as motivated by][]{2007ApJ...654..304K}, and
t$_*$ is the local free-fall timescale.

### 2013-06-25 numerical-timestep.tex
Many small changes to all files -- GB final edits

\begin{equation}
\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{\myvec{g}}} \right)_L ,
\label{eqn:dtaccel}
\end{equation}

\begin{equation}
\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{\myvec{a}_{rad}}} \right)_L,
\label{eqn:dtrad}
\end{equation}

\begin{equation}
\Delta t_{\rm cond} = \min \left(  \frac{ \kappa_{\rm cond}}{f_{\rm sp}} \frac{\Delta x^2
    n_b}{\kappa(T)} \right)_L,
\label{eqn:dtcond}
\end{equation}

@@

Equation~(\ref{eqn:dtcond}) is the stability condition for an explicit
solution to the equation of heat conduction.  In this expression,
$n_b$ is the baryon number density, 
%defined as $n_b = \rho_b / \mu
%m_p$, where $\mu$ is the mean molecular weight, 
$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
user-defined, dimensionless
conduction suppression factor whose value must be f$_{\rm sp} \leq 1$.  
%Both are defined in Section~\ref{sec.num.conductions}.  
$\kappa_{\rm cond}$ is a dimensionless prefactor whose
value must be $0 < \kappa_{\rm cond}  \leq 0.5$, and is exactly 0.5 for the
implemention in \enzo.
From a practical
perspective, it is useful to note that, unlike other timestep criteria
discussed above (which effectively scale as $\Delta x / \sqrt{T}$ with
the grid cell size $\Delta x$ and temperature $T$),
the timestep criterion due to thermal conduction scales as $\Delta x^2
/ T^{2.5}$, which can result in a rapid decrease in timestep in
regions of high resolution and/or temperature.

>>>>

\begin{equation}
\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{|\myvec{g}|}} \right)_L ,
\label{eqn:dtaccel}
\end{equation}

\begin{equation}
\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{|\myvec{a}_{rad}|}} \right)_L,
\label{eqn:dtrad}
\end{equation}

\begin{equation}
\Delta t_{\rm cond} = \min \left(  \frac{ k_{\rm cond}}{f_{\rm sp}} \frac{\Delta x^2
    n_b}{\kappa_{\rm sp}(T)} \right)_L,
\label{eqn:dtcond}
\end{equation}

@@

Equation~(\ref{eqn:dtcond}) is the stability condition for an explicit
solution to the equation of heat conduction.  In this expression,
$n_b$ is the baryon number density, 
%defined as $n_b = \rho_b / \mu
%m_p$, where $\mu$ is the mean molecular weight, 
$\kappa_{\rm sp}(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
user-defined, dimensionless
conduction suppression factor whose value must be f$_{\rm sp} \leq 1$.  
%Both are defined in Section~\ref{sec.num.conductions}.  
$k_{\rm cond}$ is a dimensionless prefactor whose
value must be $0 < k_{\rm cond}  \leq 0.5$, and is exactly 0.5 for the
implemention in \enzo.
From a practical
perspective, it is useful to note that, unlike other timestep criteria
discussed above (which effectively scale as $\Delta x / \sqrt{T}$ with
the grid cell size $\Delta x$ and temperature $T$),
the timestep criterion due to thermal conduction scales as $\Delta x^2
/ T^{2.5}$, which can result in a rapid decrease in timestep in
regions of high resolution and/or temperature.
